---
title: "Predictr Code"
author: "Kiernan Nicholls"
date: "Spring, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE, 
  message = FALSE,
  warning = FALSE,
  error   = FALSE,
  dpi = 300, 
  max.print = 50)
```

```{r install_packages, eval=FALSE}
library(devtools) # installing functions
install_cran("here") # for local storage
install_cran("tidyverse") # for data manipulation
install_cran("verification") # for forecast analysis
install_github("hrbrmstr/wayback") # for internet archives
```

```{r attach_packages}
library(readr)     # reading data
library(dplyr)     # wrangling data
library(tidyr)     # tidying data
library(stringr)   # character strings
library(wayback)   # reading archives
library(ggplot2)   # plotting data
library(magrittr)  # piping data
library(lubridate) # dates strings
```

# Read Input Data

Input data has be manually archived on the [The Wayback Machine][01] is a
digital archive of the World Wide Web run by the [Internet Archive][02], a
nonprofit organization. Using the `wayback` package, "memento" files can be
retrieved from the internet and scraped by the `readr` package into [tibble][03]
data frames.

[01]: https://archive.org/web/
[02]: https://archive.org/
[03]: https://tibble.tidyverse.org/

## Read market data

First, we will read prediction market data courtesy of [PredictIt][04], an
exchange owned and operated by the Victoria University of Wellington. As part 
of their operating agreement with the Commodity Futures Trading Commission,
PredicIt provides market history data for free to academic researchers.

The data was provided via email as a tab-separated file and can be loaded
with `readr`. Two separate files were sent with the data on the Maine 2nd and
New York 27th congressional districts, which were accidentally left out from the
the main file. All data can be found in the `/data` folder.

[04]: https://www.predictit.org/

```{r read_market}
DailyMarketData <-
  here::here("data", "DailyMarketData.csv") %>%
  read_delim(delim = "|",
             na = "n/a",
             col_types = cols(
               MarketId = col_character(),
               ContractName = col_character(),
               ContractSymbol = col_character(),
               Date = col_date(format = "")))

Market_ME02 <-
  here::here("data", "Market_ME02.csv") %>%
  read_csv(col_types = cols(ContractID = col_character(),
                            Date = col_date(format = "%m/%d/%Y")))

Contract_NY27 <-
  here::here("data" , "Contract_NY27.csv") %>%
  read_csv(na = c("n/a", "NA"),
           skip = 156,
           col_types = cols(ContractID = col_character(),
                            Date = col_date(format = "%m/%d/%Y")))

DailyMarketData
```

## Read member data

Congressional member data is used to provide party information as well as
ideology and leadership scores. The data comes from the 
[the @unitedstates project][05] and [GovTrack][06].

[05]: https://theunitedstates.io/
[06]: https://www.govtrack.us/

```{r read_members}
## Current members of the 115th
## Archived: 2018-10-22 at 18:11
legislators_current <-
  "https://theunitedstates.io/congress-legislators/legislators-current.csv" %>%
  read_memento(timestamp = "2018-10-22", as = "raw") %>%
  read_csv(col_types = cols(govtrack_id = col_character()))

# The ideology and leadership scores of the 115th
# Calculated with cosponsorship analysis
# Archived 2019-01-21 17:13:08
sponsorshipanalysis_h <-
  str_c("https://www.govtrack.us/",
        "data/analysis/by-congress/115/sponsorshipanalysis_h.txt") %>%
  read_memento(timestamp = "2019-03-23", as = "raw") %>%
  read_csv(col_types = cols(ID = col_character()))

sponsorshipanalysis_s <-
  str_c("https://www.govtrack.us/",
        "data/analysis/by-congress/115/sponsorshipanalysis_s.txt") %>%
  read_memento(timestamp = "2019-03-23", as = "raw") %>%
  read_csv(col_types = cols(ID = col_character()))

legislators_current
```

## Read model data

Forecasting model data is courtesy of [FiveThirtyEight][07], who provides the
top-level output of their proprietary model for free to the public.

[07]: https://fivethirtyeight.com/

```{r read_models}
## District level 538 House model history
## Updated:  2018-11-06 at 01:56
## Archived: 2018-11-06 at 12:06
house_district_forecast <-
  str_c(site = "https://projects.fivethirtyeight.com/",
        file = "congress-model-2018/house_district_forecast.csv") %>%
  read_memento(timestamp = "2018-11-06", as = "raw") %>%
  read_csv()

# Seat level 538 Senate model history
# Updated:  2018-11-06 at 11:06
# Archived: 2018-11-06 at 21:00
senate_seat_forecast <-
  str_c(site = "https://projects.fivethirtyeight.com/",
        file = "congress-model-2018/senate_seat_forecast.csv") %>%
  read_memento(timestamp = "2018-11-06", as = "raw") %>%
  read_csv()

house_district_forecast
```

## Read election results data

Election results data is courtesy of [FiveThirtyEight][07] and their parent
company [ABC News][08], whose Decision Desk called outcomes of races on election
night.

This data is used to assess the accuracy of each predictive method.

[08]: https://abcnews.go.com/

```{r read_results}
# Midterm election results via ABC and 538
# Used in https://53eig.ht/2PiFb0f
# Published: 2018-12-04 at 17:56
# Archived:  2018-04-04 at 16:08
forecast_results_2018 <-
  str_c(site = "https://raw.githubusercontent.com/",
        fold = "fivethirtyeight/data/master/forecast-review/",
        file = "forecast_results_2018.csv") %>%
  read_memento(timestamp = "2019-04-04", as = "raw") %>%
  read_csv(col_types  = cols(
    Democrat_Won = col_logical(),
    Republican_Won = col_logical(),
    uncalled = col_logical(),
    forecastdate = col_date(format = "%m/%d/%y"),
    category = col_factor(ordered = TRUE,
                          levels = c("Solid D",
                                     "Likely D",
                                     "Lean D",
                                     "Tossup (Tilt D)",
                                     "Tossup (Tilt R)",
                                     "Lean R",
                                     "Likely R",
                                     "Safe R"))))

forecast_results_2018
```

# Format Data for Comparison

Once data is collected from the Internet Archive, each tibble will need to
be formatted in a similar style. This will be done using `tidyverse` data 
manipulation tools.

Ultimately, each tibble will need similar `date` and `race` variables, which
together can be used to perform relational joins for comparison. Using all 4
primary data sets, we can create a tibble for each predictive method with all
the data needed for comparison.

## Format member data

```{r format_members}
members <- legislators_current %>%
  unite(first_name, last_name,
        col = name,
        sep = " ") %>%
  rename(gid     = govtrack_id,
         chamber = type,
         class   = senate_class,
         birth   = birthday) %>%
  select(name, gid, birth, state, district, class, party, gender, chamber) %>%
  arrange(chamber)

members$name     %<>% iconv(to = "ASCII//TRANSLIT")
members$name     %<>% str_replace_all("Robert Menendez", "Bob Menendez")
members$name     %<>% str_replace_all("Robert Casey",    "Bob Casey")
members$name     %<>% str_replace_all("Bernard Sanders", "Bernie Sanders")
members$chamber  %<>% recode("rep" = "house", "sen" = "senate")
members$district %<>% str_pad(width = 2, pad = "0")
members$class    %<>% str_pad(width = 2, pad = "S")
members$party    %<>% recode("Democrat"    = "D",
                             "Independent" = "D",
                             "Republican"  = "R")

members$district <- if_else(condition = is.na(members$district),
                            true = members$class,
                            false = members$district)

# Create district code as relational key
members %<>%
  unite(col = race,
        state, district,
        sep = "-",
        remove = TRUE) %>%
  select(-class) %>%
  arrange(name)

# Format member stats for join
members_stats <-
  bind_rows(sponsorshipanalysis_h, sponsorshipanalysis_s,
            .id = "chamber") %>%
  select(ID, chamber, party, ideology, leadership) %>%
  rename(gid = ID)
members_stats$chamber %<>% recode("1" = "house", "2" = "senate")
members_stats$party %<>% recode("Democrat"    = "D",
                                "Independent" = "D",
                                "Republican"  = "R")
members_stats$gid %<>% as.character()
# Add stats to frame by GovTrack ID
members %<>% inner_join(members_stats, by = c("gid", "party", "chamber"))

members
```

## Format market data

```{r foromat_markets}
markets <- DailyMarketData %>%
  rename(mid      = MarketId,
         name     = MarketName,
         symbol   = MarketSymbol,
         party    = ContractName,
         open     = OpenPrice,
         close    = ClosePrice,
         high     = HighPrice,
         low      = LowPrice,
         volume   = Volume,
         date     = Date) %>%
  select(date, everything()) %>%
  select(-ContractSymbol)

# Get candidate names from full market question
markets$name[str_which(markets$name, "Which party will")] <- NA
markets$name %<>% word(start = 2, end = 3)

# Recode party variables
markets$party %<>% recode("Democratic or DFL" = "D",
                          "Democratic"        = "D",
                          "Republican"        = "R")

# Remove year information from symbol strings
markets$symbol %<>% str_remove(".2018")
markets$symbol %<>% str_remove(".18")

# Divide the market symbol into the name and race code
markets %<>%
  separate(col = symbol,
           into = c("symbol", "race"),
           sep = "\\.",
           extra = "drop",
           fill = "left") %>%
  select(-symbol)

# Recode the original contract strings for race variables
markets$race %<>% str_replace("SENATE", "S1")
markets$race %<>% str_replace("SEN",    "S1")
markets$race %<>% str_replace("SE",     "S1")
markets$race %<>% str_replace("AL",     "01")   # at large
markets$race %<>% str_replace("OH12G",  "OH12") # not sure
markets$race %<>% str_replace("MN99",   "MNS2") # special election
markets$race[markets$name == "SPEC"] <- "MSS2"  # special election
markets$race[markets$mid  == "3857"] <- "CAS1"  # market name mustyped
markets$name[markets$name == "PARTY"] <- NA     # no name
markets$name[markets$name == "SPEC"]  <- NA     # no name

markets$race <- paste(str_sub(markets$race, 1, 2), # state abbreviation
                      sep = "-",                   # put hyphen in middle
                      str_sub(markets$race, 3, 4)) # market number)

# Remove markets incorectly repeated
# Some not running for re-election
markets %<>% filter(mid != "3455", # Paul Ryan
                    mid != "3507", # Jeff Flake
                    mid != "3539", # Shea-Porter
                    mid != "3521", # Darrell Issa
                    mid != "3522", # Repeat of 4825
                    mid != "4177", # Repeat of 4232
                    mid != "4824") # Repeat of 4776

# Divide the data based on market question syntax
# Market questions provided name or party, never both
markets_with_name <- markets %>%
  filter(is.na(party)) %>%
  select(-party)

markets_with_party <- markets %>%
  filter(is.na(name)) %>%
  select(-name)

# Join with members key to add party, then back with rest of market
markets <- markets_with_name %>%
  inner_join(members, by = c("name", "race")) %>%
  select(date, mid, race, party, open, low, high, close, volume) %>%
  bind_rows(markets_with_party)

# Add in ME-02 and NY-27 which were left out of initial data
ny_27 <- Contract_NY27 %>%
  rename_all(tolower) %>%
  slice(6:154) %>%
  mutate(mid = "4729",
         race = "NY-27",
         party = "R") %>%
  select(-average)

me_02 <- Market_ME02 %>%
  rename_all(tolower) %>%
  rename(party = longname) %>%
  filter(date != "2018-10-10") %>%
  mutate(mid = "4945",
         race = "ME-02")

markets_extra <-
  bind_rows(ny_27, me_02) %>%
  select(date, mid, race, party, open, low, high, close, volume)

markets_extra$party[str_which(markets_extra$party, "GOP")] <- "R"
markets_extra$party[str_which(markets_extra$party, "Dem")] <- "D"

# Bind with ME-02 and NY-27
markets %<>%  bind_rows(markets_extra)

markets
```

## Format model data

```{r format_model}
# Format district for race variable
model_district <- house_district_forecast %>%
  mutate(district = str_pad(string = district,
                            width = 2,
                            side = "left",
                            pad = "0"))

# Format class for race variable
model_seat <- senate_seat_forecast %>%
  rename(district = class) %>%
  mutate(district = str_pad(string = district,
                            width = 2,
                            side = "left",
                            pad = "S"))

model_combined <-
  bind_rows(model_district, model_seat, .id = "chamber") %>%
  # Create race variable for relational join
  unite(col = race,
        state, district,
        sep = "-",
        remove = TRUE) %>%
  rename(name = candidate,
         date = forecastdate,
         prob = win_probability,
         min_share = p10_voteshare,
         max_share = p90_voteshare) %>%
  filter(name != "Others") %>%
  select(date, race, name, party, chamber, everything()) %>%
  arrange(date, name)

# Recode identifying variable for clarification
model_combined$chamber %<>% recode("1" = "house",
                                   "2" = "senate")

# Only special elections are for senate.
model_combined$special[is.na(model_combined$special)] <- FALSE

# Convert percent vote share values to decimal
model_combined[, 10:12] <- model_combined[, 10:12] * 0.01

# Recode incumbent Independent senators for relational joins with Markets
# Both caucus with Democrats and were endoresed by Democratic party
model_combined$party[model_combined$name == "Bernard Sanders"]   <- "D"
model_combined$party[model_combined$name == "Angus S. King Jr."] <- "D"
model_combined %<>% filter(name != "Zak Ringelstein")

# Seperate model data by model format
# According to 538, the "classic" model can be used as a default
model <- model_combined %>% 
  filter(model == "classic") %>% 
  select(-model)

model_lite <- model_combined %>% 
  filter(model == "lite") %>% 
  select(-model)

model_deluxe <- model_combined %>% 
  filter(model == "deluxe") %>% 
  select(-model)

model
```

## Format election results

```{r format_results}
results <- forecast_results_2018 %>%
  filter(branch  != "Governor",
         version == "classic") %>%
  separate(col    = race,
           into   = c("state", "district"),
           sep    = "-") %>%
  rename(winner   = Democrat_Won) %>%
  mutate(district = str_pad(district, width = 2,  pad   = "0")) %>%
  unite(state, district,
        col = race,
        sep = "-") %>%
  select(race, winner) %>%
  filter(race != "NC-09") # Harris fraud charges

results
```

# Compare Predictive Methods

Once each data frame has been properly formatted, they can be filtered to remove
redundant predictions. Each row in both sets will contain the day's probability
of a Democratic party candidate winning.

```{r join_messy}
# Take the complimentary probability if only GOP data
# Find race codes for markets with data on only one candidate
single_party_markets <- markets %>%
  group_by(date, race) %>%
  summarise(n = n()) %>%
  filter(n == 1) %>%
  ungroup() %>%
  pull(race) %>%
  unique()

# Invert the GOP prices for markets with only GOP candidates
invert <- function(x) 1 - x

invert_gop <- markets %>%
  filter(race %in% single_party_markets,
         party == "R") %>%
  mutate(close = invert(close),
         party = "D")

# Take all but the only GOP markets
original_dem <- markets %>%
  filter(!race %in% invert_gop$race,
         party == "D")

# Combined both back together
markets2 <-
  bind_rows(original_dem, invert_gop) %>%
  select(date, race, close) %>%
  arrange(date, race)

# Create model data with only dem party info
model2 <- model %>%
  group_by(date, race, party) %>%
  summarise(prob = sum(prob)) %>%
  ungroup() %>%
  filter(party == "D") %>%
  select(-party)

# Join democratic predictions from both markets and models for comparison
# Keep market and model data in seperate columns
messy <-
  inner_join(markets2, model2,
             by  = c("date", "race")) %>%
  filter(date  >= "2018-08-01",
         date  <= "2018-11-05") %>%
  rename(model  = prob,
         market = close)

messy
```

```{r join_tidy}
# Make the data tidy with each prediction as an observation
tidy <- messy %>%
  gather(model, market,
         key = method,
         value = prob) %>%
  arrange(date, race, method)

tidy
```

```{r join_hits}
# Add in results to determine binary hits/misses
hits <- tidy %>%
  mutate(pred = prob > 0.5) %>%
  inner_join(results, by = "race") %>%
  mutate(hit = pred == winner) %>%
  select(date, race, method, prob, pred, winner, hit)

hits
```

```{r run_tests}
# Run a welch two sample t-test?
hits %$%
  t.test(formula = hit ~ method,
         alternative = "greater")

# Run a 2-sample test for equality of proportions?
hits %>%
  select(date, race, method, hit) %>%
  spread(key = method,
         value = hit) %>%
  select(market, model) %>%
  colSums() %>%
  prop.test(n = nrow(hits)/2 %>% rep(2))

hits %>%
  group_by(pred, winner, method) %>%
  summarise(prob = mean(prob),
            n = n()) %>%
  arrange(pred, winner)
```

```{r run_brier}
hits %>%
  mutate(brier_score = (winner - prob)^2) %$%
  t.test(formula = brier_score ~ method)

hits_model  <- hits %>% filter(method == "model")
hits_market <- hits %>% filter(method == "market")

brier_model <- verification::brier(
  obs = hits_model$winner,
  pred = hits_model$prob,
  baseline = rep(0.5, nrow(hits_model)),
  bins = TRUE)

brier_market <- verification::brier(
  obs = hits_market$winner,
  pred = hits_market$prob,
  baseline = rep(0.5, nrow(hits_market)),
  bins = TRUE)
```

# Explore Data Visually

```{r set_colors, echo=FALSE}
color_model  <- "#ED713A" # 538 brand color
color_market <- "#07A0BB" # PredictIt brand color
color_blue   <- "royalblue3" # Democratic
color_red    <- "red3" # Republican
```

```{r echo=FALSE}
# Distribution of original probabilities by method
# Join market onto model keep all model races
full_join(x = model, y = markets, by = c("date", "race", "party")) %>%
  # Show only 1 candidate per race
  filter(date == "2018-11-05") %>%
  select(date, race, close, prob) %>%
  rename(markets = close, model = prob) %>%
  gather(markets, model, key = method, value = prob) %>%
  mutate(method = method %>% recode("model" = "Forecasting Model",
                                    "markets" = "Prediction Markets")) %>%
  ggplot(mapping = aes(x = prob, fill = method)) +
  geom_histogram(binwidth = 0.10) +
  facet_wrap(~method, scales = "free_y", drop = TRUE) +
  scale_fill_manual(values = c(color_model, color_market)) +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.2),
                     minor_breaks = 0,
                     labels = scales::percent) +
  labs(title = "Distribution of Race Probabilities by Predictive Method",
       x = "Democratic Win Probability",
       y = "Number of Races") +
  theme(legend.position = "none")
```

```{r echo=FALSE}
# Number of dollars traded over time
markets %>%
  filter(date >= "2018-01-01", date <= "2018-11-05") %>%
  group_by(date) %>%
  mutate(traded = close * volume) %>%
  summarise(sum = sum(traded, na.rm = TRUE)) %>%
  mutate(cumsum = cumsum(sum)) %>%
  ggplot(mapping = aes(x = date, y = cumsum)) +
  geom_line(color = color_market, size = 2) +
  geom_vline(xintercept = as.Date("2018-08-01"), size = 0.5) +
  geom_vline(xintercept = as.Date("2018-11-05"), size = 0.5) +
  scale_y_continuous(labels = scales::dollar) +
  labs(title = "Cumulative Dollars Traded on Election Markets",
       x = "Date",
       y = "Dollars Traded to Date")
```

```{r echo=FALSE}
# Number of markets opened over time
markets %>%
  filter(date > "2018-01-01", date < "2018-11-05") %>%
  group_by(date) %>%
  summarise(count = n()) %>%
  ggplot(mapping = aes(x = date, y = count)) +
  geom_line(color = color_market, size = 2) +
  geom_vline(xintercept = as_date("2018-08-01"), size = 0.5) +
  geom_vline(xintercept = as_date("2018-11-05"), size = 0.5) +
  labs(title = "Cumulative Number of Election Markets",
       x = "Date",
       y = "Markets to Date")
```

```{r echo=FALSE}
# Races by Model on X and Market on Y
messy %>%
  mutate(party = "D") %>%
  filter(date == "2018-11-05") %>%
  left_join(model, by = c("date", "race", "party")) %>%
  inner_join(results, by = "race") %>%
  ggplot(aes(x  = model, y  = market)) +
  geom_hline(yintercept = 0.5) +
  geom_vline(xintercept = 0.5) +
  geom_abline(slope = 1, intercept = 0) +
  geom_label(aes(fill = winner, label = race)) +
  scale_y_continuous(labels = scales::dollar) +
  scale_x_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("red", "forestgreen")) +
  labs(title = "Midterm Races by Democrat's Chance of Winning",
       subtitle = "November 5th, Night Before Election Day",
       x = "Model Probability",
       y = "Market Price",
       shape = "Chamber",
       color = "Incumbency")
```

```{r echo=FALSE}
messy %>%
  mutate(party = "D") %>%
  filter(date == "2018-11-05") %>%
  left_join(model, by = c("date", "race", "party")) %>%
  inner_join(results, by = "race") %>%
  ggplot(aes(x  = model, y  = market)) +
  geom_hline(yintercept = 0.5) +
  geom_vline(xintercept = 0.5) +
  geom_label(mapping = aes(x = 0.25, y = 0.75, label = "Market Predicts Win"),
             label.size = 0,
             fill = "#ebebeb",
             size = 6) +
  geom_label(mapping = aes(x = 0.75, y = 0.25, label = "Model Predicts Win"),
             label.size = 0,
             fill = "#ebebeb",
             size = 6) +
  geom_label(mapping = aes(x = 0.25, y = 0.25, label = "Both Predict Loss"),
             label.size = 0,
             fill = "#ebebeb",
             size = 6) +
  geom_label(mapping = aes(x = 0.75, y = 0.75, label = "Both Predict Win"),
             label.size = 0,
             fill = "#ebebeb",
             size = 6) +
  geom_abline(slope = 1, intercept = 0, lty = 2) +
  geom_point(aes(color = winner, shape = chamber), size = 4) +
  scale_y_continuous(labels = scales::dollar) +
  scale_x_continuous(labels = scales::percent) +
  scale_color_manual(values = c("red", "forestgreen")) +
  labs(title = "Midterm Races by Democrat's Chance of Winning",
       subtitle = "November 5th, Night Before Election Day",
       x = "Model Probability",
       y = "Market Price",
       shape = "Chamber",
       color = "Democrat Won")
```

```{r echo=FALSE}
# Weird NJ-02 Market Error
markets %>%
  filter(race == "NJ-02", date > "2018-10-25") %>%
  ggplot(aes(x = date, y = close)) +
  geom_hline(yintercept = 0.5) +
  geom_line(aes(color = party), size = 2) +
  scale_color_manual(values = c(color_blue, color_red)) +
  scale_y_continuous(labels = scales::dollar) +
  scale_x_date() +
  labs(title = "Price History of New Jersey 2nd Betting Market",
       x = "Date",
       y = "Closing Price")
```

```{r echo=FALSE}
hits %>%
  mutate(month = month(date, label = TRUE)) %>%
  group_by(month, method) %>%
  summarise(prop = mean(hit, na.rm = TRUE)) %>%
  ggplot(aes(x = month, y = prop, fill = method)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c(color_market, color_model)) +
  coord_cartesian(ylim = c(0.50, 1.0)) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Proportion of Correct Predictions by Month",
       subtitle = "PredictIt Markets and FiveThirtyEight Model",
       y = "Proportion",
       x = "Month of Year")
```

```{r echo=FALSE}
hits %>%
  mutate(week = week(date)) %>%
  group_by(week, method) %>%
  summarise(prop = mean(hit, na.rm = TRUE)) %>%
  ggplot(aes(x = week, y = prop, color = method)) +
  geom_line(size = 3) +
  coord_cartesian(ylim = c(0.75, 0.95)) +
  scale_y_continuous(labels = scales::percent) +
  scale_color_manual(values = c(color_market, color_model)) +
  labs(title = "Proportion of Correct Predictions by Week",
       subtitle = "PredictIt Markets and FiveThirtyEight Model",
       y = "Proportion",
       x = "Week of Year")
```

```{r echo=FALSE}
hits %>%
  group_by(date, method) %>%
  summarise(prop = mean(hit, na.rm = TRUE)) %>%
  ggplot(aes(x = date, y = prop, color = method)) +
  geom_line(size = 2) +
  coord_cartesian(ylim = c(0.75, 0.95)) +
  scale_y_continuous(labels = scales::percent) +
  scale_color_manual(values = c(color_market, color_model)) +
  labs(title = "Proportion of Correct Predictions by Day",
       subtitle = "PredictIt Markets and FiveThirtyEight Model",
       y = "Proportion",
       x = "Day of Year")
```

```{r echo=FALSE}
hits %>%
  mutate(bin = prob %>% round(digits = 1)) %>%
  group_by(method, bin) %>%
  summarise(prop = mean(winner), n = n()) %>%
  ggplot(mapping = aes(bin, prop)) +
  geom_abline(intercept = 0, slope = 1, lty = 2)  +
  geom_point(mapping = aes(color = method, size = n), alpha = 0.75) +
  scale_x_continuous(breaks = seq(0, 1, 0.1), minor_breaks = 0,
                     labels = scales::percent) +
  scale_y_continuous(breaks = seq(0, 1, 0.1), minor_breaks = 0,
                     labels = scales::percent) +
  scale_color_manual(values = c(color_market, color_model), guide = FALSE) +
  scale_size(range = c(2, 12)) +
  theme(legend.position = "bottom", legend.key = element_blank()) +
  labs(title = "Expected Probabilities and Actual Proportions of Democratic Victory",
       subtitle = "Expected probabilities binned by rounding to the nearest 10%",
       y = "Proportion of Actual Democratic Victory",
       x = "Predicted Probability of Democratic Victory",
       size = "Number of Predictions")
```
